\bb
{dlpolymanual}
Smith, W., Forester, T.R. , ``The DlPoly 2.0 User Manual'', 
T.R.Forester, W.Smith,
CCLRC, Daresbury Laboratory, Daresbury, Warrington WA4 4AD, England (1995).

\bb
{Brainerd}
W.~S. Brainerd, C.~H. Goldberg, J.~C. Adams.
{\em Programmer's guide to Fortran 90}.
McGraw-Hill, New York, NY, 1st edition, 1990.

\bb
{Metcalf}
M.~Metcalf, J.~Reid.
{\em Fortran 90 explained}.
Oxford University Press, Oxford, UK, 1st edition, 1990.

\bb
{Kerrigan}
J.~F. Kerrigan.
{\em Migrating to Fortran 90}.
O'Reilly, Sebastopol, CA, 1st edition, 1993.

\bb
{allentildesley}
M.P.Allen e D.J.Tildesley, ``Computer simulation of liquids'',
Clarendon Press, Oxford, 1987.

\bb
{shake}
Ciccotti, G., Ryckaert, J.P., 
{\it Comp.Phys.Rep.}{\bf 4}(1986), 345

\bb
{rdshake}
Smith, W., Forester, T.R.,
{\it Comput.Phys.Commun.}{\bf 79}(1994), 63

\bb
{mttk}
Martyna, J.G., Tuckerman, M.E., Tobias, D.J., Klein, M.L.,
{\it Mol.Phys.}{\bf 87}(1996), 1117

\bb
{npt}
Melchionna, S., Ciccotti, G., Holian, B.L.,
{\it Mol.Phys.}{\bf 78}(1993), 533

\bb
{melchcicc}
Melchionna, S., Ciccotti, G., 
{\it J.Chem.Phys.}{\bf 106}(1997), 195

\bb
{marchiprocacci}
Procacci, P., Marchi, M., 
{\it J.Chem.Phys.}{\bf 104}(1996), 3003

\bb
{darden}
Essmann, U., Perera, L., Berkowitz, M.L., Darden, T., 
Lee, H., Pedersen, L.G.
{\it J.Chem.Phys.}{\bf 103}(1995), 8577

\bb
{smith-fincham}
Smith, W., Fincham, D.,
{\it Mol.Sim.}{\bf 10}(1993), 67

\bb
{Stroustrup}
B.~Stroustrup.
{\em The C++ programming language}.
Addison-Wesley, Reading, MA, 2nd edition, 1991.

\bb
{oof90_web}
C. D. Norton, V. K. Decyk, and B. K. Szymanski.
{\em High Performance Object-Oriented Programming in Fortran 90}
Internet Web Pages, Oct. 1996. http://www.cs.rpi.edu/~szymansk/oof90.html
